# EraBrewer <img align="right" src="man/figures/logo_white_bg.svg" width=400>

Palettes inspired by Taylor Swift's album covers.

Structure of the package was based on
[`MetBrewer`](https://github.com/BlakeRMills/MetBrewer/)
themselves inspired by:

* [`RColorBrewer`](https://CRAN.R-project.org/package=RColorBrewer/RColorBrewer.pdf) from [Cynthia Brewer](https://colorbrewer2.org)
* [`PNWColors`](https://github.com/jakelawlor/PNWColors)
* [`wesanderson`](https://github.com/karthik/wesanderson)

[Installation](#install-package)

[Palettes](#palettes)

The colours are just colours. The copyright for the album covers is most likely
owned by either the publisher of the work or the artist(s) which produced the
recording or cover artwork in question. It is believed that the use of
low-resolution images of such covers qualifies as fair use.

## Install Package

### `R`

Install the released version from CRAN:

```r
install.packages("EraBrewer")
```

Or the development version from GitHub:

```r
# install.packages("pak")
pak::pak("mathias-sm/EraBrewer")
```

Usage:

```r
ggplot(data=iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) +
  geom_point() +
  scale_color_manual(values=era.brewer(palette_name, n=3))
```

### `Python`

Place `Python/era_brewer.py` into your source directory.

Use it in your code:

```python
from matplotlib.colors import LinearSegmentedColormap
from cycler import cycler
import era_brewer

# Discrete
color_scheme = era_brewer.era_brew('Lover1', n=5, brew_type="discrete")
custom_cycler = cycler(color=color_scheme)

# Continuous
image_map = era_brewer.era_brew('Showgirl2',n=100,brew_type="continuous")
custom_cmap = LinearSegmentedColormap.from_list("custom_gradient", image_map)
```

## Palettes

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(cowplot)
library(tidyverse)
library(urbnmapr)
library(ggstream)
library(knitr)
library(magick)
library(devtools)
library(pkgdown)
knitr::opts_chunk$set(echo = FALSE)
devtools::load_all()   # or load_all(".")
```

```{r results='asis', fig.height=8, fig.retina = 10}
source("dev/test_plots.R")
for (palette in names(EraPalettes)) {
  cat(paste0("\n\n### ", palette, "\n\n"))
  colors = EraPalettes[palette][[1]][[1]]
  order = EraPalettes[palette][[1]][[2]]
  n = length(colors)
  path = test.plots.era(palette)
  cat(paste0("\n\n![Cover and sample plots for the album '", palette, "'](",path,")\n\n"))
  cat(paste0('\nColors: `"', paste(colors, collapse = '", "'), '"`.\n\nIf you only need $n$ colors, we suggest you use:\n\n<details>\n'))
  for (i in c(1:n)) {
    ordered <- paste(colors[which(order %in% c(1:i)==TRUE)], collapse = '", "')
    cat(paste0("\n", i, '. `"', ordered, '"`'))
  }
  cat("\n\n</details>")
}
```
